On the power of oritatami cotranscriptional folding with unary bead sequence\thanks{This work is supported in part by KAKENHI Grant-in-Aid for Challenging Research (Exploratory) No.~18K19779 and JST Program to Disseminate Tenure Tracking System No.~6F36 granted to S.~Z.~F. and S.~S.}

### 2018-12-05 tt_theorem.tex
latest
12/5 evening

\begin{proof}%[lemma \ref{TTT_entrance_Tab}]
See Fig.\ref{TTT_tunnel_direction}.
$S[i+2] \in \{ t_s, t_o \} $ means that $w[i+2]$ is stabilized by a tunnel section of type $S$ or $O$.
Thus,  its predecessor $w[i+1]$ must be inside the tunnel section, that is, $n_1$ and $n_2$ must be occupied.
Free bonds of $w[i]$, if any, cannot be used in future by another bead $w[j]$ because otherwise the part of transcript $w[i..j]$ and the bond between $w[i]$ and $w[j]$ would form a closed curve and the curve would cross the path of $C_{i-1}$ between $n_1$ and $n_2$, contradiction.
%Binding capabilities that $w[i]$ supplies are inactive according to Jordan curve theorem on $n_1$, $n_2$, and $w[i]$.
Therefore, if $w[i]$ forms a bond at its stabilization $\#bc(C_{i-1}) > \#bc(C_{i})$ holds.
We now prove that $w[i]$ must form a bond.

@@

            \draw[->, blue] (60:0.9)--(60:0.1);
            \draw[->, blue] (0:0.1)--(0:1.1);

@@

	Let us proceed to Statement 2.
	See Fig.~\ref{TTT_tunnelC_enter_usingBond}.
	Consider the case when $w[i]$ is stabilized by a tunnel section of type $S$ or $O$.
	As prove in Lemma~\ref{TTT_exit}, $\#bc(C_{i-2}) \geq \#bc(C_{i})$.%the increase of binding capability by $w[i-1]$ and $w[i]$ is at most 0.
	The bead $w[i]$ leaves two free hand, it will be used by $w[i+2]$, and hence, $w[i+2]$ does not increase the binding capability.
	Thus, $\#bc(C_{i-2}) > \#bc(C_{i+2})$.
	This argument actually work also for the case when $w[i]$ is stabilized rather by binding.

	Finally, we prove Statement 3.
	%Consider the case when $w[i]$ is stabilized by a tunnel section of type $A$ and $w[i]$ does not consumes any bond.
	In order for $w[i]$ not to bind, $w[i-2]$ must have already used up its hands.
	% The bead $w[i-2]$ must consume two bonds ; otherwise $w[i-2]$'s hands is used by $w[i]$.
	The bead $w[i-1]$ consumes one hand and provides nothing.
	Thus,  $\#bc(C_{i-3}) - 1 \geq \#bc(C_{i})$.
	The bead $w[i+1]$ consumes one hand and provides nothing.
	Finally, $w[i+2]$ uses one hand of $w[i]$, and hence, does not increase the binding capability.
	Therefore, $\#bc(C_{i-3}) -2 \geq \#bc(C_{i+2})$.
	This argument actually works also for the case when $w[i]$ is stabilized rather by binding. \qed
\end{proof}

@@

          \node[below left] at (180:1) {$w[i]$};
          \node[above] at (0:0) {$w[i+1]$};
          \node[below] at (-120:1) {$w[i+2]$};

          \node[right] at (120:1) {$n_3$};
          \node[right] at (60:1) {$n_2$};
          \node[right] at (0:1) {$n_1$};

@@

          \node[right] at (120:1) {$n_3$};
          \node[right] at (60:1) {$n_2$};
          \node[right] at (0:1) {$n_1$};

@@

          \end{scope}
        \end{tikzpicture}
        \end{minipage}
        \end{tabular}
    \caption{Case of $S[i] = t_a$ and $S[i+2] = t_a$: (Right) Moreover $w[i]$ does not consume any bond.}
    \label{TTT_tunnelC_enter_usingBond}
  \end{center}
\end{figure}

@@

	\node at (0:2.5) {$n$};
	\node[shift=(60:1)] at (0:2) {$n+1$};
    \end{tikzpicture}
    \caption{The first bead out of $\hexagon_{w[-n+1]}^n$}
    \label{TTT_a4_first}
\end{figure}

@@

%Both $n_3$ and $n_4$ are occupied because $S[i] = t$ or $t^\times$.
%Suppose $w[i]$ were stabilized without any bond.
%Then, all neighbors of $n_3$ must be occupied as long as the bead at $n_3$ is not singular; otherwise, a bead at $n_3$ would be capable of forming a bond with $w[i]$.
%That is, $n_1$ is occupied.
%We can continue this argument in clockwise manner to fill the neighbors of $w[i]$ until a singular bead is encountered, if any.
%If needed, the same argument is applied starting from $n_4$ in counter-clockwise manner.
%Thus, all neighbors of $w[i]$ must be occupied but then $S[i+1] = \blacksquare$, contradiction. \qed

@@

\begin{figure} [tb]
  \begin{center}
  \begin{tabular}{c c c}
 \begin{minipage}{0.3\hsize}
 \centering
    \begin{tikzpicture}

@@

	\node[below] at (0:0) {$w[i-1]$};
	\node[above] at (60:1) {$w[i]$};
	\node[above, shift=(0:1)] at (60:1) {$w[i+1]$};
	\end{tikzpicture}
    \end{minipage}
    \end{tabular}
    \caption{The moment when the transcript steps outside $\hexagon^n_0$.}
    \label{fig:ttt_d2}
  \end{center}
\end{figure}

@@

\begin{proof}
 We assume $w[i]$ is the first bead stabilized outside $\hexagon^n_0$.
 See Fig.~\ref{fig:ttt_d2}.
 The next bead $w[i+1]$ is to be bound for stabilization.
 Hence, it goes to the west  or to the east (Fig.~\ref{fig:ttt_d2}).
 Once $w[i+2]$ is stabilized at $p$, the remaining transcript folds into the zig-zag conformation.
 In order to avoid this or nondeterminism, $w[i+2]$ must form two bonds; it thus decreases binding capability by 1.
 Until when a bead is stabilized outside $\hexagon^{n+1}_0$, binding capability never increases because of arity being 2 and of the Tunnel Troll Theorem.
 This means that, only at most $\#bc(C_0) \leq 2n$ times we can thus expand that hexagonal region.
 In other words, outside $\hexagon^{3n}_0$ the transcript cannot help but fold into the zig-zag conformation. \qed

% 
% If $w[i+1]$ elongated to west and $w[i+2]$ forms one bond, then $w[i+2]$ is stabilized at $p_3$; otherwise $w[i+2]$ forms two bonds.
% If $w[i+1]$ elongated to west and $w[i+2]$ is stabilized at $p_3$, $w[i+1..]$ form a zig-zag conformation.
%  If $w[i+1]$ elongated to east and $w[i+2]$ forms one bond, then $w[i+2]$ is stabilized at $p_2$; otherwise $w[i+2]$ forms two bonds.
% If $w[i+1]$ elongated to east and $w[i+2]$ is stabilized at $p_2$, $w[i+1..]$ is zig-zag conformation.
% Thus, if $w[i]$ is the first bead outside $\hexagon^k_0$ for some $k \geq n$, $\#bc(C_i) > \#bc(C_{i+1})$ or $w[i+1..]$ form a zig-zag conformation.
% 
% We assume $w[l]$ is the first bead outside $\hexagon^k_0$ and $w[m]$ is the first bead outside $\hexagon^{k+1}_0$ for some $k,l,m$ with $k<3n$. Let $s$ be a element of $S[l+1..m-1]$. If $s = b$, $s$ does not increase binding capability.
% If $s=t$, $s$ is a element of a subsequence $bt^+b$. This subsequence does not increase binding capability according to Theorem~\ref{TTT}. Thus, $\#bc(C_l) > \#bc(C_{m})$ or $w[l..]$ form a zig-zag conformation.
% 
% The seed $\sigma$ provides at most $2n$ binding capability, that is, $\#bc(C_0) \leq 2n$.
% Thus,  if  $w[i]$ is the first bead outside $\hexagon^{3n}_0$, then this conformation has no hand without $w[i-1]$ and $w[i]$.
% Hence, $w[27n^2 + 9n + 1..]$ form a zig-zag conformation.

>>>>

\begin{proof}%[lemma \ref{TTT_entrance_Tab}]
See Fig.\ref{TTT_tunnel_direction}.
$S[i+2] \in \{ t_s, t_o \} $ means that $w[i+2]$ is stabilized by a tunnel section of type S or O.
Thus,  its predecessor $w[i+1]$ must be inside the tunnel section, that is, $n_1$ and $n_2$ must be occupied.
Free bonds of $w[i]$, if any, cannot be used in future by another bead $w[j]$ because otherwise the part of transcript $w[i..j]$ and the bond between $w[i]$ and $w[j]$ would form a closed curve and the curve would cross the path of $C_{i-1}$ between $n_1$ and $n_2$, contradiction.
%Binding capabilities that $w[i]$ supplies are inactive according to Jordan curve theorem on $n_1$, $n_2$, and $w[i]$.
Therefore, if $w[i]$ forms a bond at its stabilization $\#bc(C_{i-1}) > \#bc(C_{i})$ holds.
We now prove that $w[i]$ must form a bond.

@@

            \draw[->, blue] (60:0.9)--(60:0.1);
            \draw[->, blue] (0:0.1)--(0:0.9);

@@

	Let us proceed to Statement 2.
	See Fig.~\ref{TTT_tunnelC_enter_usingBond}.
	Consider the case when $w[i]$ is stabilized by a tunnel section of type S or O.
	As prove in Lemma~\ref{TTT_exit}, $\#bc(C_{i-2}) \geq \#bc(C_{i})$.	The bead $w[i]$ leaves two free hand, it will be used by $w[i+2]$, and hence, $w[i+2]$ does not increase the binding capability.
	Thus, $\#bc(C_{i-2}) > \#bc(C_{i+2})$.
	This argument actually work also for the case when $w[i]$ is stabilized rather by binding.

	Finally, we prove Statement 3.
	In order for $w[i]$ not to bind, $w[i-2]$ must have already used up its hands.
	The bead $w[i-1]$ consumes one hand and provides nothing.
	Thus,  $\#bc(C_{i-3}) - 1 \geq \#bc(C_{i})$.
	The bead $w[i+1]$ consumes one hand and provides nothing.
	Finally, $w[i+2]$ uses one hand of $w[i]$, and hence, does not increase the binding capability.
	Therefore, $\#bc(C_{i-3}) -2 \geq \#bc(C_{i+2})$.
	This argument actually works also for the case when $w[i]$ is stabilized rather by binding. \qed
\end{proof}

@@

          \node[below left] at (180:1) {$w[i]$};
          \node[above] at (0:0) {$w[i+1]$};
          \node[below] at (-120:1) {$w[i+2]$};
          \begin{scope}[shift=(-120:1)]
            \fill[blue](0,0) circle [radius=0.05];

@@

@@

          \end{scope}
        \end{tikzpicture}
        \end{minipage}
        \end{tabular}
    \caption{The bead $w[i+2]$ is stabilized by a tunnel of type A. (Right) Moreover $S[i] = t_a$.}
    \label{TTT_tunnelC_enter_usingBond}
  \end{center}
\end{figure}

@@

	\node at (0:2.5) {$n$};
	\node[shift=(60:1)] at (0:2) {$n+1$};
    \end{tikzpicture}
    \caption{The first bead out of $\hexagon_{O}^n$}
    \label{TTT_a4_first}
\end{figure}

@@

@@

\begin{figure} [tb]
  \begin{center}
  \begin{tabular}{c c c}
 \begin{minipage}{0.3\hsize}
 \centering
    \begin{tikzpicture}
      \draw[white] (0:0)-- (60:2);

@@

	\node[below] at (0:0) {$w[i-1]$};
	\node[above] at (60:1) {$w[i]$};
	\node[above, shift=(0:1)] at (60:1) {$w[i+1]$};
	\end{tikzpicture}
    \end{minipage}
    \end{tabular}
    \caption{The moment when the transcript steps outside $\hexagon^n_O$.}
    \label{fig:ttt_d2}
  \end{center}
\end{figure}

@@

\begin{proof}
 We assume $w[i]$ is the first bead stabilized outside $\hexagon^n_O$.
 See Fig.~\ref{fig:ttt_d2}.
 The next bead $w[i+1]$ is to be bound for stabilization.
 Hence, it goes to the west  or to the east (Fig.~\ref{fig:ttt_d2}).
 Once $w[i+2]$ is stabilized at $p$, the remaining transcript folds into the zig-zag conformation.
 In order to avoid this or nondeterminism, $w[i+2]$ must form two bonds; it thus decreases binding capability by 1.
 Until when a bead is stabilized outside $\hexagon^{n+1}_O$, binding capability never increases because of arity being 2 and of the Tunnel Troll Theorem.
 This means that, only at most $\#bc(C_0) \leq 2n$ times we can thus expand that hexagonal region.
 In other words, outside $\hexagon^{3n}_O$ the transcript cannot help but fold into the zig-zag conformation. \qed
